`r opts_chunk$set(cache=FALSE, fig.width=10, message=FALSE, warning=FALSE)`
<link href="http://joey711.github.com/phyloseq/markdown.css" rel="stylesheet"></link>

# [Waste Not, Want Not Supplemental Files](http://joey711.github.io/waste-not-supplemental/)

This repo hosts the supplemental source/output files corresponding to **Protocol 1** of the following article.

#### McMurdie and Holmes (2014) Waste Not, Want Not: Why Rarefying Microbiome Data is Inadmissible. *PLoS Computational Biology*
(in press)

Please make sure to cite the above article if you make use of these supplemental files.

---

# Source Code Index

## Quick Links to Output 

- [Source Code Index](source-code-index.html)
- [Minimal Rarefy Example](minimal-rarefy-example/minimal-rarefy-example.html)
- [Simulations Design](simulation-design/simulation-design-figure-ex-matrix.html)
- [Dispersion Survey](dispersion-survey/dispersion.html)
- [Cluster Accuracy Simulation](simulation-cluster-accuracy/simulation-cluster-accuracy-server.html)
- [Differential Abundance Simulation](simulation-differential-abundance/simulation-differential-abundance-server.html)

## The R Markdown Source Code

For more advanced users interested in seeing and/or modifying the [R markdown](http://www.rstudio.com/ide/docs/r_markdown) source code from which these reproducible, web browser-friendly output files were created.

```{r find-parse-source-files, echo=FALSE, include=FALSE}
files = list.files(pattern="\\.[Rr]md", recursive=TRUE)
files = files[!files=="source-code-index.Rmd"]
```

```{r rmd-links, results='asis', echo=FALSE}
rmdlines = paste("### [", files, "](", files, ")", sep="")
cat(rmdlines, sep="\n\n")
```

---

# Additional Links

## [Report an Issue](https://github.com/joey711/waste-not-supplemental/issues)

## [GitHub Repository Branch](https://github.com/joey711/waste-not-supplemental/tree/gh-pages)

## [Phyloseq Extensions](http://joey711.github.io/phyloseq-extensions/)

For brief reproducible examples of related extensions for the phyloseq package, see the [phyloseq extensions front page](http://joey711.github.io/phyloseq-extensions/).